---
title: "HW2"
author: "Hakan Ekiz - IE48B - Fall2021"
date: "22 Kasım 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# HW2

## Introduction

In this study, we will work with the GBF Training dataset. With the 1D Fused method, we will work on the Regression method and then compare them. In the 1D Fused method, we will try to establish the best model by finding the best lambda values for the time series in our dataset. In the regression tree method, we will find the optimal "maxdepth" number.



```{r, include=FALSE}
require(data.table)
require(ggplot2)
require(repr)
require(rpart)
require(rattle)
require(TSrepr)
require(zoo)
require(genlasso)
```

## Data

First of all, we take our GBFTrain dataset as a data table and then we make the necessary manipulations to make it suitable for timeseries studies.

```{r}

options(repr.plot.width=15, repr.plot.height=8)
current_folder=getwd()
dataset='CBF'

train_data_path=sprintf('%s/CBFData/%s_TRAIN.txt',current_folder,dataset)
train_data=fread(train_data_path)


```
 "C:/Users/Z0047JBE/Desktop/DERSLER SON/IE48B/HW2"

```{r}
setnames(train_data,'V1','class')
train_data=train_data[order(class)]
train_data[,class:=as.character(class)]
train_data[,id:=1:.N]

```


```{r}
long_train=melt(train_data,id.vars=c('id','class'))
#head(long_train)
long_train[,time:=as.numeric(gsub("\\D", "", variable))-1]
#long_train
long_train=long_train[,list(id,class,time,value)]
long_train=long_train[order(id,time)]
head(long_train)


```


We have 2 data sets that we will use for modeling and then use for comparisons, we will continue with the methods after plotting the data set we have once and looking at the situation.



```{r , echo=FALSE}

ggplot(long_train, aes(time,value)) + geom_line(aes(color=as.character(id))) +
     facet_wrap(~class)

```


# 1D Fused Lasso


We will use this package by downloading the genlasso package for the 1D Fused Lasso method. Here we will use k=10 for the penalty value. Let's do a test for our values with id=1 before applying it to the entire dataset.


```{r}

out1 = fusedlasso1d(as.numeric(train_data[id==1, 2:129]))

cv1 = cv.trendfilter(out1, k = 10)

lambda = cv1$lambda
error = cv1$err
df = data.table(lambda,error)
df[,id:=1:.N]
df
```


To check lambda values and min error values, let's do the same for values with id=2.

```{r}
out2 = fusedlasso1d(as.numeric(train_data[id==2, 2:129]))

cv2 = cv.trendfilter(out2, k = 10)

lambda = cv2$lambda
error = cv2$err
df2 = data.table(lambda,error)
df2[,id:=1:.N]
df2
```



```{r}
a = cv1$lambda.min
plot(out1, lambda = a)

```


After the necessary evaluations and trials, we now have to apply this for all id values with a for loop.



```{r}
out = vector()
lamb = vector()
fit = vector()



for (i in c(1:30)) {
  outt = fusedlasso1d(as.numeric(train_data[id==i, 2:129]))
  out[i]= outt
  cvv = cv.trendfilter(outt, k = 10)
  lamb[i] = cvv$lambda.min
  fit[(128*i-127):(128*i)] = outt$fit[,match(lamb[i], outt$lambda)]
  
  
  #c = cvv$lambda.min
  #plot(outt, lambda=c)

}

```



```{r}
dfall = data.table(lamb)
dfall[,id:=1:.N]
dfall

```



# Reg Tree

As the other method, we will apply the regression tree method. Here we will use minsplit=20, minbucket=10 and cp=0. We need to find the maxdepth number and the optimal maxdepth numbers for our values and create trees accordingly.

```{r}
selected_dt=long_train[id==1]

tree1=rpart(value~time,selected_dt,control=rpart.control(minsplit=20, minbucket = 10, cp=0, maxdepth=4))

fancyRpartPlot(tree1)
```

```{r}

tree1$cptable

```



```{r}
require(e1071)

depth = vector()
pred = vector()

ranges <- list(minsplit=20, cp=0, minbucket=10, maxdepth=1:30)


for (i in 1:30) {
  deep <- tune(rpart, value~time, data=long_train[id==i], ranges=ranges)

  depth[i] = deep$best.parameters[,4]
  
  treett = rpart(value~time, long_train[id==i], control=rpart.control(minsplit=20, minbucket=10, cp=0, maxdepth=depth[i]))
  predi = predict(treett,long=train[id==i])
  pred[(128*i-127):(128*i)] = predi
}

```


```{r}

dfdepth = data.table(depth)
dfdepth[,id:=1:.N]
dfdepth


```


We currently know the optimal maxdepth numbers for each id, so we can also inspect one by one by creating trees for each value we want to see. 
Example for id=16:


```{r}
selected_dt=long_train[id==16]

tree16=rpart(value~time,selected_dt,control=rpart.control(minsplit=20, minbucket = 10, cp=0, maxdepth=5))

fancyRpartPlot(tree16)
```


```{r}
selected_dt[,tree_rep:=predict(tree16,selected_dt)]

data_plot=melt(selected_dt,id.vars='time',measure.vars=c('value','tree_rep'))

ggplot(data_plot,aes(x=time,y=value,color=variable))+
    geom_line()
```

# Comp


We will compare the estimates we made with the above two methods with the real values by noticing them.

```{r}
long_train[, lasso := t(fit)]
long_train[, tree:= t(pred)]
long_train
```

```{r}
msetree = vector()
mselasso = vector()

for (i in 1:30) {
  
  msetree[i] <- mean((long_train[id==i]$value - long_train[id==i]$tree)^2)
  mselasso[i] <- mean((long_train[id==i]$value - long_train[id==i]$lasso)^2)
  
}


```



```{r}

dfcomp = data.table(mselasso,msetree)
dfcomp[,id:=1:.N]
dfcomp

```

When we examine the MSE values in the table, we can see that the estimations we made with the 1D Fused Lasso method have lower MSE values than the Regression Tree. As a result, we can say that we generally make closer estimations with the 1D Fused Lasso method.










