---
title: "HW3-Plane"
author: "Hakan Ekiz - IE48B - Fall2021"
date: "13 Aralık 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Homework Information

### Representations

1- Raw
2- Difference
3- Piecewise Aggregate Approximation (segmenth length=12)
4- Piecewise Lnear Approximation
5- Tree (maxdepth=12)

### Distance Measurements

1- Euclidian
2- DTW (Windows size=10)
3- DTW (Window size=20)
4- ERP

### K nn classifiers

1- k=1
2- k=3
3- k=5

### Datasets

1- ECG200
2- Plane
3- SyntheticControl
4- Trace
5- SmoothSubspace

Note: Since it takes time to run 5 different datasets in the same file and convert them to reports, 5 different files will be loaded for 5 different datasets.



## Libraries and Functions

```{r, message=FALSE}
require(data.table)
require(ggplot2)
require(repr)
require(rpart)
require(rattle)
require(TSrepr)
require(TSdist)
require(dtw)
require(zoo)
require(genlasso)
require(e1071)
```

```{r}
options(repr.plot.width=15, repr.plot.height=8)
```


```{r}
nn_classify_cv=function(dist_matrix,train_class,test_indices,k=1){
    
    test_distances_to_train=dist_matrix[test_indices,]
    test_distances_to_train=test_distances_to_train[,-test_indices]
    train_class=train_class[-test_indices]
    #print(str(test_distances_to_train))
    ordered_indices=apply(test_distances_to_train,1,order)
    if(k==1){
        nearest_class=as.numeric(trainclass[as.numeric(ordered_indices[1,])])
        nearest_class=data.table(id=test_indices,nearest_class)
    } else {
        nearest_class=apply(ordered_indices[1:k,],2,function(x) {trainclass[x]})
        nearest_class=data.table(id=test_indices,t(nearest_class))
    }
    
    long_nn_class=melt(nearest_class,'id')

    class_counts=long_nn_class[,.N,list(id,value)]
    class_counts[,predicted_prob:=N/k]
    wide_class_prob_predictions=dcast(class_counts,id~value,value.var='predicted_prob')
    wide_class_prob_predictions[is.na(wide_class_prob_predictions)]=0
    class_predictions=class_counts[,list(predicted=value[which.max(N)]),by=list(id)]
    
    
    return(list(prediction=class_predictions,prob_estimates=wide_class_prob_predictions))
    
}
```

# Plane

## Data 

Train Size=105
TestSize=105
Length=144
Number of Class=7


```{r}
current_folder=getwd()
dataset='Plane'
main_path=sprintf('%s/ClassificationData/%s/%s',current_folder,dataset,dataset)

dist_path=sprintf('%s/ClassificationData/%s/distances/%s',current_folder,dataset,dataset)

train_data_path=sprintf('%s_TRAIN.txt',main_path)
traindata=as.matrix(fread(train_data_path))

test_data_path=sprintf('%s_TEST.txt',main_path)
testdata=as.matrix(fread(test_data_path))

#first column is the class variable
trainclass=traindata[,1] # takes -1 and 1
testclass=testdata[,1] # takes -1 and 1

#drop first column
traindata=traindata[,2:ncol(traindata)]
testdata=testdata[,2:ncol(testdata)]

# instance characteristics
tlength=ncol(traindata)
n_series_train=nrow(traindata)
n_series_test=nrow(testdata)
```


```{r}
head(data.table(traindata))

```


```{r}
dt_ts_train=data.table(traindata)
dt_ts_train[,id:=1:.N]
long_train=melt(dt_ts_train,id.vars=c('id'))
long_train[,time:=as.numeric(gsub("\\D", "", variable))-1]
long_train=long_train[order(id,time)]
long_train[,diff_series:=value-shift(value,1),by=list(id)]
head(long_train)
head(dt_ts_train)

```

```{r}

dt_ts_test=data.table(testdata)
dt_ts_test[,id:=1:.N]
long_test=melt(dt_ts_test,id.vars=c('id'))
long_test[,time:=as.numeric(gsub("\\D", "", variable))-1]
long_test=long_test[order(id,time)]
long_test[,diff_series:=value-shift(value,1),by=list(id)]
head(long_test)
head(dt_ts_test)
```


```{r}
ggplot(long_train, aes(time,value)) + geom_line(aes(color=as.character(id))) 
```

## Representations


PAA with segmenth length = 5
PLA with n_important = 10
Tree with maxdepth = 12

```{r}
#Piecewise Aggregate + Piecewise Linear Approximation + Tree


segmentlength = 5
n_important = 10

long_train=long_train[order(id,time)]

treeres <- vector('list', n_series_train)
paares <- vector('list', n_series_train)

for (i in 1:n_series_train) {
  
  data_plot=long_train[id==i]
  
  data_ts=long_train[id==i]$value
  paa_rep=repr_paa(data_ts, segmentlength , meanC)
  dummy_time = c(1:(length(paa_rep)-1))*segmentlength
  dummy_time = c(dummy_time, (nrow(data_plot)))
  dt_paa = data.table(time=dummy_time, paa_rep)
  
  pla_rep=repr_pla(data_ts, times = n_important, return = "points")
  pla_rep=repr_pla(data_ts, times = n_important, return = "both")
  setnames(pla_rep,c('places','points'),c('time','pla_rep'))
  pla_rep$time=pla_rep$time+1
  
  
  tree=rpart(value~time,data_plot,control=rpart.control(cp=0,maxdepth=12,minsplit=20,minbucket=10))
  data_plot[,tree_rep:=predict(tree,data_plot)]
  
  
  data_plot=merge(data_plot,dt_paa,by='time',all.x=T)
  data_plot=merge(data_plot,pla_rep,by='time',all.x=T)
  
  data_plot[,paa_rep:=nafill(paa_rep,'nocb')]
  data_plot[, pla_rep := na.approx(pla_rep)]
  
  paares[[i]]=data_plot

}
results <- rbindlist(paares) 

```


```{r}

head(results)

```

Sample representaion for id=10

```{r}

selected=results[id==10]
data_show=melt(selected,id.vars='time',measure.vars=c('value','tree_rep', 'paa_rep', 'pla_rep'))
ggplot(data_show,aes(x=time,y=value,color=variable))+
    geom_line()

```


## Distance Measurement

Preliminary work with results dataset for future distance calculations;

Euc
Dtw (window size = 10)
Dtw (window size = 20)
ERP

```{r}
value_train=dcast(results[!is.na(value)],id~time,value.var='value')
head(value_train)
value_train=as.matrix(value_train)
```


```{r}
diff_train=dcast(results[!is.na(diff_series)],id~time,value.var='diff_series')
head(diff_train)
diff_train=as.matrix(diff_train)
```


```{r}
tree_train=dcast(results[!is.na(tree_rep)],id~time,value.var='tree_rep')
head(tree_train)
tree_train=as.matrix(tree_train)
```

```{r}
paa_train=dcast(results[!is.na(paa_rep)],id~time,value.var='paa_rep')
head(paa_train)
paa_train=as.matrix(paa_train)
```

```{r}
pla_train=dcast(results[!is.na(pla_rep)],id~time,value.var='pla_rep')
head(pla_train)
pla_train=as.matrix(pla_train)
```

```{r}
large_number=10000
```

### Euclidian

Raw, difference, tree, paa(piecewise aggregate approximation with segmenth length 5) and pla values are used. 

```{r}
# raw + euc
dist_euc=as.matrix(dist(value_train))
diag(dist_euc)=large_number
fwrite(dist_euc,sprintf('%s_euc_raw_dist.csv',dist_path),col.names=F)

# diif + euc
dist_diff=as.matrix(dist(diff_train))
diag(dist_diff)=large_number
fwrite(dist_diff,sprintf('%s_euc_diff_dist.csv',dist_path),col.names=F)

# tree + euc
dist_tree=as.matrix(dist(tree_train))
diag(dist_tree)=large_number
fwrite(dist_tree,sprintf('%s_euc_tree_dist.csv',dist_path),col.names=F)

# paa + euc
dist_paa=as.matrix(dist(paa_train))
diag(dist_paa)=large_number
fwrite(dist_paa,sprintf('%s_euc_paa_dist.csv',dist_path),col.names=F)

# pla + euc
dist_pla=as.matrix(dist(pla_train))
diag(dist_pla)=large_number
fwrite(dist_pla,sprintf('%s_euc_pla_dist.csv',dist_path),col.names=F)

```


```{r}

# raw + DTW(Window size = 10)
dist_dtw=as.matrix(dtwDist(value_train, window.type='sakoechiba',window.size=10))
diag(dist_dtw)=large_number
fwrite(dist_dtw,sprintf('%s_dtw_raw_dist_10ws.csv',dist_path),col.names=F)

# diff + DTW(Window size = 10)
dist_dtw_diff=as.matrix(dtwDist(diff_train, window.type='sakoechiba',window.size=10))
diag(dist_dtw_diff)=large_number
fwrite(dist_dtw_diff,sprintf('%s_dtw_diff_dist_10ws.csv',dist_path),col.names=F)

# tree + DTW(window size = 10)
dist_dtw_tree=as.matrix(dtwDist(tree_train, window.type='sakoechiba',window.size=10))
diag(dist_dtw_tree)=large_number
fwrite(dist_dtw_tree,sprintf('%s_dtw_tree_dist_10ws.csv',dist_path),col.names=F)

# paa + DTW(window size = 10)
dist_dtw_paa=as.matrix(dtwDist(paa_train, window.type='sakoechiba',window.size=10))
diag(dist_dtw_paa)=large_number
fwrite(dist_dtw_paa,sprintf('%s_dtw_paa_dist_10ws.csv',dist_path),col.names=F)

# pla + DTW(window size = 10)
dist_dtw_pla=as.matrix(dtwDist(pla_train, window.type='sakoechiba',window.size=10))
diag(dist_dtw_pla)=large_number
fwrite(dist_dtw_pla,sprintf('%s_dtw_pla_dist_10ws.csv',dist_path),col.names=F)



```


```{r}
# raw + DTW(Window size = 20)
dist_dtw2=as.matrix(dtwDist(value_train, window.type='sakoechiba',window.size=20))
diag(dist_dtw2)=large_number
fwrite(dist_dtw2,sprintf('%s_dtw_raw_dist_20ws.csv',dist_path),col.names=F)

# diff + DTW(Window size = 20)
dist_dtw_diff2=as.matrix(dtwDist(diff_train, window.type='sakoechiba',window.size=20))
diag(dist_dtw_diff2)=large_number
fwrite(dist_dtw_diff2,sprintf('%s_dtw_diff_dist_20ws.csv',dist_path),col.names=F)

# tree + DTW(window size = 20)
dist_dtw_tree2=as.matrix(dtwDist(tree_train, window.type='sakoechiba',window.size=20))
diag(dist_dtw_tree2)=large_number
fwrite(dist_dtw_tree2,sprintf('%s_dtw_tree_dist_20ws.csv',dist_path),col.names=F)

# paa + DTW(window size = 20)
dist_dtw_paa2=as.matrix(dtwDist(paa_train, window.type='sakoechiba',window.size=20))
diag(dist_dtw_paa2)=large_number
fwrite(dist_dtw_paa2,sprintf('%s_dtw_paa_dist_20ws.csv',dist_path),col.names=F)

# pla + DTW(window size = 20)
dist_dtw_pla2=as.matrix(dtwDist(pla_train, window.type='sakoechiba',window.size=20))
diag(dist_dtw_pla2)=large_number
fwrite(dist_dtw_pla2,sprintf('%s_dtw_pla_dist_20ws.csv',dist_path),col.names=F)
```

```{r}

# raw + ERP(g=1)
dist_erp=TSDatabaseDistances(value_train,distance='erp',g=1)
dist_erp=as.matrix(dist_erp)
diag(dist_erp)=large_number
fwrite(dist_erp,sprintf('%s_erp_raw_gap_1.csv',dist_path),col.names=F)

# diff + ERP(g=1)
dist_erp_diff=TSDatabaseDistances(diff_train,distance='erp',g=1)
dist_erp_diff=as.matrix(dist_erp_diff)
diag(dist_erp_diff)=large_number
fwrite(dist_erp_diff,sprintf('%s_erp_diff_gap_1.csv',dist_path),col.names=F)

# tree + ERP(g=1)
dist_erp_tree=TSDatabaseDistances(tree_train,distance='erp',g=1)
dist_erp_tree=as.matrix(dist_erp_tree)
diag(dist_erp_tree)=large_number
fwrite(dist_erp_tree,sprintf('%s_erp_tree_gap_1.csv',dist_path),col.names=F)

# paa + ERP(g=1)
dist_erp_paa=TSDatabaseDistances(paa_train,distance='erp',g=1)
dist_erp_paa=as.matrix(dist_erp_paa)
diag(dist_erp_paa)=large_number
fwrite(dist_erp_paa,sprintf('%s_erp_paa_gap_1.csv',dist_path),col.names=F)

# pla + ERP(g=1)
dist_erp_pla=TSDatabaseDistances(pla_train,distance='erp',g=1)
dist_erp_pla=as.matrix(dist_erp_pla)
diag(dist_erp_pla)=large_number
fwrite(dist_erp_pla,sprintf('%s_erp_pla_gap_1.csv',dist_path),col.names=F)


```

## Comparisons 

```{r}
# cv indices start here
require(TunePareto)

set.seed(13429)
nof_rep=5
n_fold=10
cv_indices=generateCVRuns(trainclass, ntimes =nof_rep, nfold = n_fold, 
                          leaveOneOut = FALSE, stratified = TRUE)

str(cv_indices)

dist_folder=sprintf('%s/ClassificationData/%s/distances/',current_folder,dataset)
dist_files=list.files(dist_folder, full.names=T)
```

```{r}
list.files(dist_folder)
```

```{r}
k_levels=c(1,3,5)
approach_file=list.files(dist_folder)
result=vector('list',length(dist_files)*nof_rep*n_fold*length(k_levels))
iter=1
for(m in 1:length(dist_files)){ #
    print(dist_files[m])
    dist_mat=as.matrix(fread(dist_files[m],header=FALSE))
    for(i in 1:nof_rep){
        this_fold=cv_indices[[i]]
        for(j in 1:n_fold){
            test_indices=this_fold[[j]]
            for(k in 1:length(k_levels)){
                current_k=k_levels[k]
                current_fold=nn_classify_cv(dist_mat,trainclass,test_indices,k=current_k)
                accuracy=sum(trainclass[test_indices]==current_fold$prediction$predicted)/length(test_indices)
                tmp=data.table(approach=approach_file[m],repid=i,foldid=j,
                               k=current_k,acc=accuracy)
                result[[iter]]=tmp
                iter=iter+1
                
            }
            
        }
    
    }   
    
}
```

```{r}
overall_results=rbindlist(result)
summarized_results=overall_results[,list(avg_acc=mean(acc),sdev_acc=sd(acc),result_count=.N),by=list(approach,k)]
summarized_results[order(-avg_acc)]

```

```{r}
summarized_results[which.max(summarized_results$avg_acc)]
```

Dynamic Time Warping method (window size=20) with difference representation and k=5 gives best results between my trials.


```{r}
require(ggplot2)
ggplot(overall_results,aes(x=paste0(approach,'+',k), y=acc)) +
geom_boxplot()+
coord_flip()
```


