---
title: "Project"
author: "Hakan Ekiz - IE48B - Fall2021"
date: "5 Ocak 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Introduction 

In this project, we will try to estimate the difference in trading data in the Turkish electricity market, that is, the imbalance values on a daily basis.In doing so, we will use two different datasets: Bulk-imbalance and Weather files. 
After developing the model, we will try to predict the hourly imbalance values for the next  day's 12 hours using the daily data.
Actually, what we need to do is to find system direction.
If the bulk imbalance value is less than -50, system direction is negative;if between -50 and +50, it will be neutral; If it is more than +50, it will be positive.

In our project, as a step, 
1) To adapt and combine 2 different datasets for modelling 
2) Creating a simple time series model 
3) Develop the model (with EBLR) 
4) get the results next day.




```{r, message=FALSE}
require(data.table)
require(ggplot2)
require(repr)
require(rpart)
require(rattle)
require(TSrepr)
require(TSdist)
require(dtw)
require(zoo)
require(genlasso)
require(e1071)
require(lubridate)
require(forecast)
require(dplyr)
```

## Data


```{r}
options(repr.plot.width=15, repr.plot.height=8)

current_folder=getwd()
```

We have 2 different datasets. Bulk_imbalance and weather, first we should import and examni them. 

```{r}

dat=fread('bulk_imbalance2.csv')
weat=fread('2022-01-18_weather.csv')
head(dat)
head(weat)

```


Because we will not use ZeroCoded and OneCoded valeus we eliminate these values. After that, because we should combine 2 dataet, add datetime (Year,month,day, hour) column into the dataset.

```{r}

c1 = dat
c1 = c1[,list(date,hour,net,system_direction)]
c1[,datetime:=ymd(date)+dhours(hour)]
c1=c1[order(datetime)]

c2 = c1[,list(datetime,net,system_direction)]
head(c2)


```

In the weather dataset we have 7 different variable values for 7 variable. Firstly, we should simplify the lat,lon and variable values. After we have simpler dataset, we should change its form with dcast. 

```{r}
w1 = weat
w1[,datetime:=ymd(date)+dhours(hour)]
w1=w1[order(datetime)]

w1 = w1 %>% 
  mutate(loc = 
           case_when(lat <= 36.50 ~ "loc1" ,
                     lat==37.00 ~ "loc2" ,
                     lat==38.00 ~ "loc3" ,
                     lat==38.50 ~ "loc4" ,
                     lat==39.75 ~ "loc5" ,
                     lat==40.00 ~ "loc6" ,
                     lat>=41.00 ~ "loc7" )
         )
w1= as.data.table(w1)
w1[,class:=paste(loc,variable)]

w2 = w1[,list(datetime,class,value)]
w2[,class:=as.factor(class)]


head(w2)

```


```{r}
trainw2=dcast(w2[!is.na(value)],datetime~class,value.var='value')
names(trainw2) <- gsub(" ", "_", names(trainw2))

head(trainw2)
```

Now we have 2 datasets whic are suitable to be used in models and be arranged with each other. 


```{r}

all = merge(c2,trainw2, by="datetime")
head(all)
```


```{r}
ggplot(all ,aes(x=datetime,y=net)) + geom_line()
```

In the combined dataset, we reveal the values date, weekday, hour, and month using the datetime value.
After all these operations, the dataset we get will be like this:

```{r}
all[,trnd:=1:.N]

all[,w_day:=as.character(wday(datetime,label=T))]
all[,mon:=as.character(month(datetime,label=T))]
all[,hour:=as.factor(hour(datetime))]
all[,Date:=as.Date(datetime)]
head(all)
```

## Models

### Simple Model1 


First of all, we wanted to see what kind of result it would give if we put all the variables in the dataset we obtained into a linear model.


```{r}

lm1=lm(net~trnd+w_day+mon+hour+
         loc1_DSWRF_surface+loc1_RH_2.m.above.ground+loc1_TCDC_low.cloud.layer+loc1_TMP_2.m.above.ground+loc1_wdir_10m+loc1_ws_10m+
         loc2_DSWRF_surface+loc2_RH_2.m.above.ground+loc2_TCDC_low.cloud.layer+loc2_TMP_2.m.above.ground+loc2_wdir_10m+loc1_ws_10m+
         loc3_DSWRF_surface+loc3_RH_2.m.above.ground+loc3_TCDC_low.cloud.layer+loc3_TMP_2.m.above.ground+loc3_wdir_10m+loc1_ws_10m+
         loc4_DSWRF_surface+loc4_RH_2.m.above.ground+loc4_TCDC_low.cloud.layer+loc4_TMP_2.m.above.ground+loc4_wdir_10m+loc1_ws_10m+
         loc5_DSWRF_surface+loc5_RH_2.m.above.ground+loc5_TCDC_low.cloud.layer+loc5_TMP_2.m.above.ground+loc5_wdir_10m+loc1_ws_10m+
         loc6_DSWRF_surface+loc6_RH_2.m.above.ground+loc6_TCDC_low.cloud.layer+loc6_TMP_2.m.above.ground+loc6_wdir_10m+loc1_ws_10m+
         loc7_DSWRF_surface+loc7_RH_2.m.above.ground+loc7_TCDC_low.cloud.layer+loc7_TMP_2.m.above.ground+loc7_wdir_10m+loc1_ws_10m
       ,all)


summary(lm1)


```


```{r}
tmp=copy(all)
tmp[,actual:=net]
tmp[,predicted:=predict(lm1,tmp)]
#head(tmp)
ggplot(tmp[Date=='2019-05-08'] ,aes(x=datetime)) +
        geom_line(aes(y=actual,color='real')) + 
        geom_line(aes(y=predicted,color='predicted'))
```

Since it does not give a good result, we will start to create a simpler model and develop it with the EBLR method.


### Model2


In this model, we start by using only the values we infer from time


```{r}
lm2=lm(net~trnd+w_day+mon+hour,all)
summary(lm2)
```

```{r}
tmp2=copy(all)
tmp2[,actual:=net]
tmp2[,predicted:=predict(lm2,tmp2)]
#head(tmp)
ggplot(tmp2[Date=='2019-05-08'] ,aes(x=datetime)) +
        geom_line(aes(y=actual,color='real')) + 
        geom_line(aes(y=predicted,color='predicted'))
```

```{r}
checkresiduals(lm2$residuals)

```

We check the residual values of this model because we are trying to determine which values in the past it relates to. 
After examining, we add these values as variables, as we see that the correlation is high in the values of 24 lag (1 day before) and 168 lag (1 week before).

```{r}
tmp2[,prevday:=shift(x=tmp2$net,n=24L,fill=mean(tmp2$net))]
tmp2[,prevweek:=shift(x=tmp2$net,n=168L,fill=mean(tmp2$net))]
head(tmp2)
```

### Improvements


First, we add prevday and prevweek values to our previous model. Then, we examine the residuals of this new model on the tree.

```{r}
lm2_1=lm(net~trnd+w_day+mon+hour+prevday+prevweek,
                 tmp2)
summary(lm2_1)
```

```{r}
tmp2[,predicted:=predict(lm2_1,tmp2)]
#head(tmp)
ggplot(tmp2[Date=='2019-05-08'] ,aes(x=datetime)) +
        geom_line(aes(y=actual,color='real')) + 
        geom_line(aes(y=predicted,color='predicted'))
```



```{r}


tmp2[,residual:=actual-predicted]


fit_res_tree=rpart(residual~w_day+mon+hour+prevday+prevweek+
         loc1_DSWRF_surface+loc1_RH_2.m.above.ground+loc1_TCDC_low.cloud.layer+loc1_TMP_2.m.above.ground+loc1_wdir_10m+loc1_ws_10m+
         loc2_DSWRF_surface+loc2_RH_2.m.above.ground+loc2_TCDC_low.cloud.layer+loc2_TMP_2.m.above.ground+loc2_wdir_10m+loc2_ws_10m+
         loc3_DSWRF_surface+loc3_RH_2.m.above.ground+loc3_TCDC_low.cloud.layer+loc3_TMP_2.m.above.ground+loc3_wdir_10m+loc3_ws_10m+
         loc4_DSWRF_surface+loc4_RH_2.m.above.ground+loc4_TCDC_low.cloud.layer+loc4_TMP_2.m.above.ground+loc4_wdir_10m+loc4_ws_10m+
         loc5_DSWRF_surface+loc5_RH_2.m.above.ground+loc5_TCDC_low.cloud.layer+loc5_TMP_2.m.above.ground+loc5_wdir_10m+loc5_ws_10m+
         loc6_DSWRF_surface+loc6_RH_2.m.above.ground+loc6_TCDC_low.cloud.layer+loc6_TMP_2.m.above.ground+loc6_wdir_10m+loc6_ws_10m+
         loc7_DSWRF_surface+loc7_RH_2.m.above.ground+loc7_TCDC_low.cloud.layer+loc7_TMP_2.m.above.ground+loc7_wdir_10m+loc7_ws_10m,tmp2,
                   control=rpart.control(cp=0,maxdepth=3))

fancyRpartPlot(fit_res_tree)

```

As seen in the tree, the rightmost path causes many errors. That's why we focus on this part first.And we add our newly created decision variables to the model.Thus, we obtain a new model with better results. Then, we re-examine the residual values of this new model by creating a tree.

We repeat this process until our improvement is very low.
The process:
-Create a model
-Examine the tree for residuals
-Create decision variables cause many errors
-Add the decision varianles to model and re-evaluate the new model


```{r}

tmp2[,loc7tmp2ag:=as.numeric(loc7_TMP_2.m.above.ground>=30)]
tmp2[,pv:=as.numeric(prevweek>=2070)]
tmp2[,loc7tmp2ag2:=as.numeric(loc7_TMP_2.m.above.ground>=31)]

lm2_2=lm(net~trnd+w_day+mon+hour+prevday+prevweek+loc7tmp2ag:pv:loc7tmp2ag2,
                 tmp2)
summary(lm2_2)



```


```{r}

tmp2[,predicted:=predict(lm2_2,tmp2)]
#head(tmp)
ggplot(tmp2[Date=='2019-05-08'] ,aes(x=datetime)) +
        geom_line(aes(y=actual,color='real')) + 
        geom_line(aes(y=predicted,color='predicted'))
```


```{r}


tmp2[,residual:=actual-predicted]


fit_res_tree=rpart(residual~w_day+mon+hour+prevday+prevweek+
         loc1_DSWRF_surface+loc1_RH_2.m.above.ground+loc1_TCDC_low.cloud.layer+loc1_TMP_2.m.above.ground+loc1_wdir_10m+loc1_ws_10m+
         loc2_DSWRF_surface+loc2_RH_2.m.above.ground+loc2_TCDC_low.cloud.layer+loc2_TMP_2.m.above.ground+loc2_wdir_10m+loc2_ws_10m+
         loc3_DSWRF_surface+loc3_RH_2.m.above.ground+loc3_TCDC_low.cloud.layer+loc3_TMP_2.m.above.ground+loc3_wdir_10m+loc3_ws_10m+
         loc4_DSWRF_surface+loc4_RH_2.m.above.ground+loc4_TCDC_low.cloud.layer+loc4_TMP_2.m.above.ground+loc4_wdir_10m+loc4_ws_10m+
         loc5_DSWRF_surface+loc5_RH_2.m.above.ground+loc5_TCDC_low.cloud.layer+loc5_TMP_2.m.above.ground+loc5_wdir_10m+loc5_ws_10m+
         loc6_DSWRF_surface+loc6_RH_2.m.above.ground+loc6_TCDC_low.cloud.layer+loc6_TMP_2.m.above.ground+loc6_wdir_10m+loc6_ws_10m+
         loc7_DSWRF_surface+loc7_RH_2.m.above.ground+loc7_TCDC_low.cloud.layer+loc7_wdir_10m+loc7_ws_10m,tmp2,
                   control=rpart.control(cp=0,maxdepth=3))

fancyRpartPlot(fit_res_tree)

```


```{r}

tmp2[,loc5tmp2ag:=as.numeric(loc5_TMP_2.m.above.ground>=29)]
tmp2[,pd:=as.numeric(prevday>=2329)]
tmp2[,workday:=as.numeric(!w_day %in% c('Cmt','Paz'))]

lm2_3=lm(net~trnd+w_day+mon+hour+prevday+prevweek+loc7tmp2ag:pv:loc7tmp2ag2+loc5tmp2ag:pd:workday,
                 tmp2)
summary(lm2_3)



```



```{r}

tmp2[,predicted:=predict(lm2_3,tmp2)]
#head(tmp)
ggplot(tmp2[Date=='2019-05-08'] ,aes(x=datetime)) +
        geom_line(aes(y=actual,color='real')) + 
        geom_line(aes(y=predicted,color='predicted'))
```


```{r}
tmp2[,residual:=actual-predicted]


fit_res_tree=rpart(residual~mon+hour+
         loc1_DSWRF_surface+loc1_RH_2.m.above.ground+loc1_TCDC_low.cloud.layer+loc1_TMP_2.m.above.ground+loc1_wdir_10m+loc1_ws_10m+
         loc2_DSWRF_surface+loc2_RH_2.m.above.ground+loc2_TCDC_low.cloud.layer+loc2_TMP_2.m.above.ground+loc2_wdir_10m+loc2_ws_10m+
         loc3_DSWRF_surface+loc3_RH_2.m.above.ground+loc3_TCDC_low.cloud.layer+loc3_TMP_2.m.above.ground+loc3_wdir_10m+loc3_ws_10m+
         loc4_DSWRF_surface+loc4_RH_2.m.above.ground+loc4_TCDC_low.cloud.layer+loc4_TMP_2.m.above.ground+loc4_wdir_10m+loc4_ws_10m+
         loc5_DSWRF_surface+loc5_RH_2.m.above.ground+loc5_TCDC_low.cloud.layer+loc5_wdir_10m+loc5_ws_10m+
         loc6_DSWRF_surface+loc6_RH_2.m.above.ground+loc6_TCDC_low.cloud.layer+loc6_TMP_2.m.above.ground+loc6_wdir_10m+loc6_ws_10m+
         loc7_DSWRF_surface+loc7_RH_2.m.above.ground+loc7_TCDC_low.cloud.layer+loc7_wdir_10m+loc7_ws_10m,tmp2,
                   control=rpart.control(cp=0,maxdepth=3))

fancyRpartPlot(fit_res_tree)

```


```{r}

tmp2[,loc4tmp2ag:=as.numeric(loc4_TMP_2.m.above.ground<8.6)]
tmp2[,whour:=as.numeric(as.numeric(hour)<9 | as.numeric(hour)>18)]
tmp2[,loc3wdir:=as.numeric(loc3_wdir_10m<336)]

lm2_4=lm(net~trnd+w_day+mon+hour+prevday+prevweek+loc7tmp2ag:pv:loc7tmp2ag2+loc5tmp2ag:pd:workday+loc4tmp2ag:whour:loc3wdir,
                 tmp2)
summary(lm2_4)



```



```{r}

tmp2[,predicted:=predict(lm2_4,tmp2)]
#head(tmp)
ggplot(tmp2[Date=='2019-05-08'] ,aes(x=datetime)) +
        geom_line(aes(y=actual,color='real')) + 
        geom_line(aes(y=predicted,color='predicted'))
```


```{r}

tmp2[,residual:=actual-predicted]


fit_res_tree=rpart(residual~mon+
         loc1_DSWRF_surface+loc1_RH_2.m.above.ground+loc1_TCDC_low.cloud.layer+loc1_TMP_2.m.above.ground+loc1_wdir_10m+loc1_ws_10m+
         loc2_DSWRF_surface+loc2_RH_2.m.above.ground+loc2_TCDC_low.cloud.layer+loc2_TMP_2.m.above.ground+loc2_wdir_10m+loc2_ws_10m+
         loc3_DSWRF_surface+loc3_RH_2.m.above.ground+loc3_TCDC_low.cloud.layer+loc3_TMP_2.m.above.ground+loc3_ws_10m+
         loc4_DSWRF_surface+loc4_RH_2.m.above.ground+loc4_TCDC_low.cloud.layer+loc4_wdir_10m+loc4_ws_10m+
         loc5_DSWRF_surface+loc5_RH_2.m.above.ground+loc5_TCDC_low.cloud.layer+loc5_wdir_10m+loc5_ws_10m+
         loc6_DSWRF_surface+loc6_RH_2.m.above.ground+loc6_TCDC_low.cloud.layer+loc6_TMP_2.m.above.ground+loc6_wdir_10m+loc6_ws_10m+
         loc7_DSWRF_surface+loc7_RH_2.m.above.ground+loc7_TCDC_low.cloud.layer+loc7_wdir_10m+loc7_ws_10m,tmp2,
                   control=rpart.control(cp=0,maxdepth=3))

fancyRpartPlot(fit_res_tree)

```

```{r}

tmp2[,loc6tmp2ag:=as.numeric(loc6_TMP_2.m.above.ground>=26)]
tmp2[,loc7ws:=as.numeric(loc7_ws_10m<3)]
tmp2[,loc1wdir:=as.numeric(loc1_wdir_10m>=172)]

lm2_5=lm(net~trnd+w_day+mon+hour+prevday+prevweek+loc7tmp2ag:pv:loc7tmp2ag2+loc5tmp2ag:pd:workday+loc4tmp2ag:whour:loc3wdir+loc6tmp2ag:loc7ws:loc1wdir,
                 tmp2)
summary(lm2_5)


```

```{r}
tmp2[,predicted:=predict(lm2_5,tmp2)]
#head(tmp)
ggplot(tmp2[Date=='2019-05-08'] ,aes(x=datetime)) +
        geom_line(aes(y=actual,color='real')) + 
        geom_line(aes(y=predicted,color='predicted'))
```


```{r}

tmp2[,residual:=actual-predicted]


fit_res_tree=rpart(residual~mon+
         loc1_DSWRF_surface+loc1_RH_2.m.above.ground+loc1_TCDC_low.cloud.layer+loc1_TMP_2.m.above.ground+loc1_ws_10m+
         loc2_DSWRF_surface+loc2_RH_2.m.above.ground+loc2_TCDC_low.cloud.layer+loc2_TMP_2.m.above.ground+loc2_wdir_10m+loc2_ws_10m+
         loc3_DSWRF_surface+loc3_RH_2.m.above.ground+loc3_TCDC_low.cloud.layer+loc3_TMP_2.m.above.ground+loc3_ws_10m+
         loc4_DSWRF_surface+loc4_RH_2.m.above.ground+loc4_TCDC_low.cloud.layer+loc4_wdir_10m+loc4_ws_10m+
         loc5_DSWRF_surface+loc5_RH_2.m.above.ground+loc5_TCDC_low.cloud.layer+loc5_wdir_10m+loc5_ws_10m+
         loc6_DSWRF_surface+loc6_RH_2.m.above.ground+loc6_TCDC_low.cloud.layer+loc6_wdir_10m+loc6_ws_10m+
         loc7_DSWRF_surface+loc7_RH_2.m.above.ground+loc7_TCDC_low.cloud.layer+loc7_wdir_10m,tmp2,
                   control=rpart.control(cp=0,maxdepth=3))

fancyRpartPlot(fit_res_tree)

```

```{r}

tmp2[,loc4wdir:=as.numeric(loc4_wdir_10m<170)]
tmp2[,loc7wdir:=as.numeric(loc7_wdir_10m>=214)]
tmp2[,loc3tmp2ag:=as.numeric(loc3_TMP_2.m.above.ground>=20)]

lm2_6=lm(net~trnd+w_day+mon+hour+prevday+prevweek+loc7tmp2ag:pv:loc7tmp2ag2+loc5tmp2ag:pd:workday+loc4tmp2ag:whour:loc3wdir+loc6tmp2ag:loc7ws:loc1wdir
         +loc4wdir:loc7wdir+loc3tmp2ag,
                 tmp2)
summary(lm2_6)


```

```{r}
tmp2[,predicted:=predict(lm2_6,tmp2)]
#head(tmp)
ggplot(tmp2[Date=='2019-05-08'] ,aes(x=datetime)) +
        geom_line(aes(y=actual,color='real')) + 
        geom_line(aes(y=predicted,color='predicted'))
```

```{r}

tmp2[,residual:=actual-predicted]


fit_res_tree=rpart(residual~mon+
         loc1_DSWRF_surface+loc1_RH_2.m.above.ground+loc1_TCDC_low.cloud.layer+loc1_TMP_2.m.above.ground+loc1_ws_10m+
         loc2_DSWRF_surface+loc2_RH_2.m.above.ground+loc2_TCDC_low.cloud.layer+loc2_TMP_2.m.above.ground+loc2_wdir_10m+loc2_ws_10m+
         loc3_DSWRF_surface+loc3_RH_2.m.above.ground+loc3_TCDC_low.cloud.layer+loc3_ws_10m+
         loc4_DSWRF_surface+loc4_RH_2.m.above.ground+loc4_TCDC_low.cloud.layer+loc4_ws_10m+
         loc5_DSWRF_surface+loc5_RH_2.m.above.ground+loc5_TCDC_low.cloud.layer+loc5_wdir_10m+loc5_ws_10m+
         loc6_DSWRF_surface+loc6_RH_2.m.above.ground+loc6_TCDC_low.cloud.layer+loc6_wdir_10m+loc6_ws_10m+
         loc7_DSWRF_surface+loc7_RH_2.m.above.ground+loc7_TCDC_low.cloud.layer,tmp2,
                   control=rpart.control(cp=0,maxdepth=3))

fancyRpartPlot(fit_res_tree)

```



```{r}

tmp2[,loc3tcdc:=as.numeric(loc3_TCDC_low.cloud.layer>=8.6)]
tmp2[,mon1:=as.numeric(!mon %in% c('Haz','Tem','Ağu','Eyl','Eki'))]
tmp2[,loc4rh:=as.numeric(loc4_RH_2.m.above.ground<69)]

lm2_7=lm(net~trnd+w_day+mon+hour+prevday+prevweek+loc7tmp2ag:pv:loc7tmp2ag2+loc5tmp2ag:pd:workday+loc4tmp2ag:whour:loc3wdir+loc6tmp2ag:loc7ws:loc1wdir
         +loc4wdir:loc7wdir+loc3tmp2ag+loc3tcdc:mon1:loc4rh,
                 tmp2)
summary(lm2_7)


```

```{r}
tmp2[,predicted:=predict(lm2_7,tmp2)]
#head(tmp)
ggplot(tmp2[Date=='2019-05-08'] ,aes(x=datetime)) +
        geom_line(aes(y=actual,color='real')) + 
        geom_line(aes(y=predicted,color='predicted'))
```


```{r}

tmp2[,residual:=actual-predicted]


fit_res_tree=rpart(residual~
         loc1_DSWRF_surface+loc1_RH_2.m.above.ground+loc1_TCDC_low.cloud.layer+loc1_TMP_2.m.above.ground+loc1_ws_10m+
         loc2_DSWRF_surface+loc2_RH_2.m.above.ground+loc2_TCDC_low.cloud.layer+loc2_TMP_2.m.above.ground+loc2_wdir_10m+loc2_ws_10m+
         loc3_DSWRF_surface+loc3_RH_2.m.above.ground+loc3_ws_10m+
         loc4_DSWRF_surface+loc4_TCDC_low.cloud.layer+loc4_ws_10m+
         loc5_DSWRF_surface+loc5_RH_2.m.above.ground+loc5_TCDC_low.cloud.layer+loc5_wdir_10m+loc5_ws_10m+
         loc6_DSWRF_surface+loc6_RH_2.m.above.ground+loc6_TCDC_low.cloud.layer+loc6_wdir_10m+loc6_ws_10m+
         loc7_DSWRF_surface+loc7_RH_2.m.above.ground+loc7_TCDC_low.cloud.layer,tmp2,
                   control=rpart.control(cp=0,maxdepth=3))

fancyRpartPlot(fit_res_tree)

```


```{r}

tmp2[,loc4ws:=as.numeric(loc4_ws_10m<3.5)]
tmp2[,loc6rh:=as.numeric(loc6_RH_2.m.above.ground<40)]
tmp2[,loc1tmp2ag:=as.numeric(loc1_TMP_2.m.above.ground>=15)]

lm2_8=lm(net~trnd+w_day+mon+hour+prevday+prevweek+loc7tmp2ag:pv:loc7tmp2ag2+loc5tmp2ag:pd:workday+loc4tmp2ag:whour:loc3wdir+loc6tmp2ag:loc7ws:loc1wdir
         +loc4wdir:loc7wdir+loc3tmp2ag+loc3tcdc:mon1:loc4rh+loc4ws:loc6rh:loc1tmp2ag,
                 tmp2)
summary(lm2_8)


```

```{r}
tmp2[,predicted:=predict(lm2_8,tmp2)]
#head(tmp)
ggplot(tmp2[Date=='2019-05-08'] ,aes(x=datetime)) +
        geom_line(aes(y=actual,color='real')) + 
        geom_line(aes(y=predicted,color='predicted'))
```
```{r}

tmp2[,residual:=actual-predicted]


fit_res_tree=rpart(residual~
         loc1_DSWRF_surface+loc1_RH_2.m.above.ground+loc1_TCDC_low.cloud.layer+loc1_ws_10m+
         loc2_DSWRF_surface+loc2_RH_2.m.above.ground+loc2_TCDC_low.cloud.layer+loc2_TMP_2.m.above.ground+loc2_wdir_10m+loc2_ws_10m+
         loc3_DSWRF_surface+loc3_RH_2.m.above.ground+loc3_ws_10m+
         loc4_DSWRF_surface+loc4_TCDC_low.cloud.layer+
         loc5_DSWRF_surface+loc5_RH_2.m.above.ground+loc5_TCDC_low.cloud.layer+loc5_wdir_10m+loc5_ws_10m+
         loc6_DSWRF_surface+loc6_TCDC_low.cloud.layer+loc6_wdir_10m+loc6_ws_10m+
         loc7_DSWRF_surface+loc7_RH_2.m.above.ground+loc7_TCDC_low.cloud.layer,tmp2,
                   control=rpart.control(cp=0,maxdepth=3))

fancyRpartPlot(fit_res_tree)

```


```{r}

tmp2[,loc5ws:=as.numeric(loc5_ws_10m<2.4)]
tmp2[,loc2tmp2ag:=as.numeric(loc2_TMP_2.m.above.ground<2.3)]
tmp2[,loc7rh2:=as.numeric(loc7_RH_2.m.above.ground<63)]

lm2_9=lm(net~trnd+w_day+mon+hour+prevday+prevweek+loc7tmp2ag:pv:loc7tmp2ag2+loc5tmp2ag:pd:workday+loc4tmp2ag:whour:loc3wdir+loc6tmp2ag:loc7ws:loc1wdir
         +loc4wdir:loc7wdir+loc3tmp2ag+loc3tcdc:mon1:loc4rh+loc4ws:loc6rh:loc1tmp2ag+loc5ws:loc2tmp2ag:loc7rh2,
                 tmp2)
summary(lm2_9)


```

```{r}
tmp2[,predicted:=predict(lm2_9,tmp2)]
#head(tmp)
ggplot(tmp2[Date=='2019-05-08'] ,aes(x=datetime)) +
        geom_line(aes(y=actual,color='real')) + 
        geom_line(aes(y=predicted,color='predicted'))
```


```{r}

tmp2[,residual:=actual-predicted]


fit_res_tree=rpart(residual~
         loc1_DSWRF_surface+loc1_RH_2.m.above.ground+loc1_TCDC_low.cloud.layer+loc1_ws_10m+
         loc2_DSWRF_surface+loc2_RH_2.m.above.ground+loc2_TCDC_low.cloud.layer+loc2_wdir_10m+loc2_ws_10m+
         loc3_DSWRF_surface+loc3_RH_2.m.above.ground+loc3_ws_10m+
         loc4_DSWRF_surface+loc4_TCDC_low.cloud.layer+
         loc5_DSWRF_surface+loc5_RH_2.m.above.ground+loc5_TCDC_low.cloud.layer+loc5_wdir_10m+
         loc6_DSWRF_surface+loc6_TCDC_low.cloud.layer+loc6_wdir_10m+loc6_ws_10m+
         loc7_DSWRF_surface+loc7_TCDC_low.cloud.layer,tmp2,
                   control=rpart.control(cp=0,maxdepth=3))

fancyRpartPlot(fit_res_tree)

```


```{r}

tmp2[,loc5rh:=as.numeric(loc5_RH_2.m.above.ground<16)]
tmp2[,loc1ws:=as.numeric(loc1_ws_10m>=3.9)]
tmp2[,loc2dswrf:=as.numeric(loc2_DSWRF_surface<841)]

lm2_10=lm(net~trnd+w_day+mon+hour+prevday+prevweek+loc7tmp2ag:pv:loc7tmp2ag2+loc5tmp2ag:pd:workday+loc4tmp2ag:whour:loc3wdir+loc6tmp2ag:loc7ws:loc1wdir
         +loc4wdir:loc7wdir+loc3tmp2ag+loc3tcdc:mon1:loc4rh+loc4ws:loc6rh:loc1tmp2ag+loc5ws:loc2tmp2ag:loc7rh2+loc5rh:loc1ws:loc2dswrf,
                 tmp2)
summary(lm2_10)




```

```{r}
tmp2[,predicted:=predict(lm2_10,tmp2)]
#head(tmp)
ggplot(tmp2[Date=='2019-05-08'] ,aes(x=datetime)) +
        geom_line(aes(y=actual,color='real')) + 
        geom_line(aes(y=predicted,color='predicted'))
```


```{r}

tmp2[,residual:=actual-predicted]


fit_res_tree=rpart(residual~
         loc1_DSWRF_surface+loc1_RH_2.m.above.ground+loc1_TCDC_low.cloud.layer+
         loc2_RH_2.m.above.ground+loc2_TCDC_low.cloud.layer+loc2_wdir_10m+loc2_ws_10m+
         loc3_DSWRF_surface+loc3_RH_2.m.above.ground+loc3_ws_10m+
         loc4_DSWRF_surface+loc4_TCDC_low.cloud.layer+
         loc5_DSWRF_surface+loc5_TCDC_low.cloud.layer+loc5_wdir_10m+
         loc6_DSWRF_surface+loc6_TCDC_low.cloud.layer+loc6_wdir_10m+loc6_ws_10m+
         loc7_DSWRF_surface+loc7_TCDC_low.cloud.layer,tmp2,
                   control=rpart.control(cp=0,maxdepth=3))

fancyRpartPlot(fit_res_tree)

```

```{r}

tmp2[,loc6tcdc:=as.numeric(loc6_TCDC_low.cloud.layer>=84)]
tmp2[,loc1dswrf:=as.numeric(loc1_DSWRF_surface>=90)]
tmp2[,loc2rh2:=as.numeric(loc2_RH_2.m.above.ground>=91)]

lm2_11=lm(net~trnd+w_day+mon+hour+prevday+prevweek+loc7tmp2ag:pv:loc7tmp2ag2+loc5tmp2ag:pd:workday+loc4tmp2ag:whour:loc3wdir+loc6tmp2ag:loc7ws:loc1wdir
         +loc4wdir:loc7wdir+loc3tmp2ag+loc3tcdc:mon1:loc4rh+loc4ws:loc6rh:loc1tmp2ag+loc5ws:loc2tmp2ag:loc7rh2+loc5rh:loc1ws:loc2dswrf
         +loc6tcdc:loc1dswrf:loc2rh2,
                 tmp2)
summary(lm2_11)


```

```{r}
tmp2[,predicted:=predict(lm2_11,tmp2)]
#head(tmp)
ggplot(tmp2[Date=='2019-05-08'] ,aes(x=datetime)) +
        geom_line(aes(y=actual,color='real')) + 
        geom_line(aes(y=predicted,color='predicted'))
```


```{r}

tmp2[,residual:=actual-predicted]


fit_res_tree=rpart(residual~
         loc1_RH_2.m.above.ground+loc1_TCDC_low.cloud.layer+
         loc2_TCDC_low.cloud.layer+loc2_wdir_10m+loc2_ws_10m+
         loc3_DSWRF_surface+loc3_RH_2.m.above.ground+loc3_ws_10m+
         loc4_DSWRF_surface+loc4_TCDC_low.cloud.layer+
         loc5_DSWRF_surface+loc5_TCDC_low.cloud.layer+loc5_wdir_10m+
         loc6_DSWRF_surface+loc6_wdir_10m+loc6_ws_10m+
         loc7_DSWRF_surface+loc7_TCDC_low.cloud.layer,tmp2,
                   control=rpart.control(cp=0,maxdepth=3))

fancyRpartPlot(fit_res_tree)

```

```{r}

tmp2[,loc3ws:=as.numeric(loc3_ws_10m<1.1)]
tmp2[,loc1tcdc:=as.numeric(loc1_TCDC_low.cloud.layer>=63)]
tmp2[,loc5wdir:=as.numeric(loc5_wdir_10m<69)]

lm2_12=lm(net~trnd+w_day+mon+hour+prevday+prevweek+loc7tmp2ag:pv:loc7tmp2ag2+loc5tmp2ag:pd:workday+loc4tmp2ag:whour:loc3wdir+loc6tmp2ag:loc7ws:loc1wdir
         +loc4wdir:loc7wdir+loc3tmp2ag+loc3tcdc:mon1:loc4rh+loc4ws:loc6rh:loc1tmp2ag+loc5ws:loc2tmp2ag:loc7rh2+loc5rh:loc1ws:loc2dswrf
         +loc6tcdc:loc1dswrf:loc2rh2+loc3ws:loc1tcdc:loc5wdir,
                 tmp2)
summary(lm2_12)


```

```{r}
tmp2[,predicted:=predict(lm2_12,tmp2)]
#head(tmp)
ggplot(tmp2[Date=='2019-05-08'] ,aes(x=datetime)) +
        geom_line(aes(y=actual,color='real')) + 
        geom_line(aes(y=predicted,color='predicted'))




```

```{r}
tmp2[,residual:=actual-predicted]


fit_res_tree=rpart(residual~
         loc1_RH_2.m.above.ground+
         loc2_TCDC_low.cloud.layer+loc2_wdir_10m+loc2_ws_10m+
         loc3_DSWRF_surface+loc3_RH_2.m.above.ground+
         loc4_DSWRF_surface+loc4_TCDC_low.cloud.layer+
         loc5_DSWRF_surface+loc5_TCDC_low.cloud.layer+
         loc6_DSWRF_surface+loc6_wdir_10m+loc6_ws_10m+
         loc7_DSWRF_surface+loc7_TCDC_low.cloud.layer,tmp2,
                   control=rpart.control(cp=0,maxdepth=3))

fancyRpartPlot(fit_res_tree)
```


```{r}
tmp2[,loc3rh:=as.numeric(loc3_RH_2.m.above.ground<25)]
tmp2[,loc5tcdc:=as.numeric(loc5_TCDC_low.cloud.layer<0.15)]
tmp2[,loc1rh:=as.numeric(loc1_RH_2.m.above.ground>=15)]

lm2_13=lm(net~trnd+w_day+mon+hour+prevday+prevweek+loc7tmp2ag:pv:loc7tmp2ag2+loc5tmp2ag:pd:workday+loc4tmp2ag:whour:loc3wdir+loc6tmp2ag:loc7ws:loc1wdir
         +loc4wdir:loc7wdir+loc3tmp2ag+loc3tcdc:mon1:loc4rh+loc4ws:loc6rh:loc1tmp2ag+loc5ws:loc2tmp2ag:loc7rh2+loc5rh:loc1ws:loc2dswrf
         +loc6tcdc:loc1dswrf:loc2rh2+loc3ws:loc1tcdc:loc5wdir+loc3rh:loc5tcdc:loc1rh,
                 tmp2)
summary(lm2_13)

tmp2[,predicted:=predict(lm2_13,tmp2)]

```


```{r}
checkresiduals(lm2_13$residuals)

```

```{r}

tmp2[,residuals13 := residuals(lm2_13)]
acf(residuals(lm2_13) , lag.max = 160)


```

```{r}
ggplot(tmp2[Date=='2019-05-08'] ,aes(x=datetime)) +
        geom_line(aes(y=actual,color='real')) + 
        geom_line(aes(y=predicted,color='predicted'))

```


### Submission


As mentioned before, we need to submit by estimating the system direction data, not the net bulk imbalance values. Therefore, we need to create system direction values from the imbalance values we estimated.



```{r}
tmp2 = tmp2 %>% 
  mutate(predicted_diretion = 
           case_when(predicted < -50 ~ "Negative" ,
                     predicted>=-50 & predicted<=50 ~ "Neutral" ,
                     predicted> 50 ~ "Positive" )
         )
tmp2= as.data.table(tmp2)
tmp2
```

```{r}
submission = tmp2[,list(Date,hour,system_direction,predicted_diretion)]
tail(submission,12)
```






